---
title: "MP_SRF_18S_amplicons_tax"
author: "lrubinat"
date: "05/03/2016"
output:
  html_document:
    theme: united
    toc: yes
  pdf_document:
    highlight: zenburn
    toc: yes
---

<!--- INITIALIZATION
```{r, echo=FALSE}
#error hook to kill knitr in case of errors
library(knitr)
knit_hooks$set(error = function(x, options) stop(x))
opts_chunk$set(cache=TRUE, autodep=TRUE)
```
--->

# 1) Data overview

Samples with less than 5000 reads are removed: 

``` {r load_data, echo=FALSE, message=FALSE}
#setwd("/genwork/lrubinat")
setwd("/home/laura/Documents/TFM/genwork/data_analysis/MP_16S_SRF_amplicons/MP_16S_amplicons_tax/")

#read data 
otu_tb16_tax <- read.table(file="/home/laura/Documents/TFM/home/data/MALASPINA/Malaspina_16S_Surface/table_with_BLAST/MP_16S_SRF_tax_classif_v2.txt", head=TRUE, fill=TRUE)

#table dimensions and format before setting column names
dim(otu_tb16_tax) # 1178  125
otu_tb16_tax[1:5,1:5]

#row names = OTU name (option A)
row.names(otu_tb16_tax)<-otu_tb16_tax[,1]

#row names = row number (option B)
#rownames(otu_tb18) <- 1:nrow(otu_tb18)

otu_tb16_tax<-otu_tb16_tax[,-1]
otu_tb16_tax[is.na(otu_tb16_tax)]<-0

dim(otu_tb16_tax)
otu_tb16_tax[1:5,1:5]

#table with occurence data alone
otu_tb16_tax_occur <- otu_tb16_tax[, 1:124]
dim(otu_tb16_tax_occur) #  56116   124
otu_tb16_tax_occur[1:5,1:5]

amplicons_per_sample_tb16<-colSums(otu_tb16_tax_occur)
amplicons_per_sample_tb16[which(colSums(otu_tb16_tax_occur)<5862)]

#remove samples with less than 5862 reads
otu_tax_occur_min5862 <- otu_tb16_tax_occur[,colSums(otu_tb16_tax_occur) > 5861]

# samples st122, st124, st137 and st144 have less than 5862 reads.


#remove samples with less than 5000 reads in the MP_18S_ss5000 dataset (so that we can compare the relative abundance of 16S and 18S OTUs considering the same samples)
#otu_tb16_min5000_v2<-subset(otu_tb16_min5000, select=-c(MD985))
```

Table dimensions and content outline:

```{r starting_dataset, echo=FALSE}
dim(otu_tax_occur_min5862)
otu_tax_occur_min5862[1:5,1:5]
```

Minimum number of reads per station:

```{r reads_per_sample_overview1, echo=1}
min(colSums(otu_tax_occur_min5862)) 
#5862
```

Maximum number of reads per station:

```{r reads_per_sample_overview2, echo=1}
max(colSums(otu_tax_occur_min5862)) 
# max: 156843
```

Identification of station with higher number of reads:

```{r reads_per_sample_overview3, echo=TRUE}
amplicons_per_sample<-colSums(otu_tax_occur_min5862)
amplicons_per_sample[which(colSums(otu_tax_occur_min5862)>155000)]
```

Overall reads per sample:

``` {r reads_per_sample_overview4, echo=FALSE}
plot(sort(colSums(otu_tax_occur_min5862)), pch=19, xlab="sample", ylab="reads per sample", cex=0.9)
```


# 2) Normalization to 5862 reads per sample

Let's normalize the original dataset by randomly subsampling 5862 reads in each station:

``` {r species_richness_rarefaction1, echo=TRUE}
library(vegan)
otu_tax_occur_min5862_t<-t(otu_tax_occur_min5862)
otu_tb16_occur_ss5862<-rrarefy(otu_tax_occur_min5862_t, 5862)
```

The normalized table shows the following dimensions and format:

```{r species_richness_rarefaction2, echo=FALSE}
dim(otu_tb16_occur_ss5862)
otu_tb16_occur_ss5862[1:5,1:5]
```

Its content fits with the expected normalization values (5000 reads per station):

``` {r species_richness_rarefaction3, echo=TRUE}
rowSums(otu_tb16_occur_ss5862)
```

Let's check out how many OTUs don't appear in the new table:

```{r species_richness_rarefaction4, echo=1:5}
length(which(colSums(otu_tb16_occur_ss5862)==0)) 
```

There are 138 OTUs that don't show any occurrence in the normalized data. Let's remove them from the table and take a look at its final dimensions:

```{r species_richness_rarefaction5, echo=1:3}
otu_tb16_occur_ss5862_no_cero<-otu_tb16_occur_ss5862[,-(which(colSums(otu_tb16_occur_ss5862)==0))]
dim(otu_tb16_occur_ss5862_no_cero)

#the final dimensions of the normalized table are 120 1177.
```

Datasets summary:
otu_tb16_tax --> 124 1315
otu_tb16_occur_ss5862_no_cero --> 119 1177



```{r merge_tables, echo=FALSE}
otu_tb_with_tax<-merge(otu_tb16_tax, otu_tb16_occur_ss5862_no_cero, by="row.names")

rownames.otu_tb_with_tax<-otu_tb_with_tax[,1]
otu_tb_with_tax<-otu_tb_with_tax[,-1]
colnames(otu_tb_with_tax, do.NULL=F)

dim(otu_tb_with_tax)
otu_tb_with_tax[1:5, 1:5]
```
